\forestclaw: Hybrid forest-of-octrees AMR for hyperbolic conservation laws

### 2013-07-31 docs/parco13/conc.tex
Make figures fit, improvise conclusion

We have presented the integration of an MPI-based forest-of-octrees adaptive
meshing code, \pforest, with a numerical solver for hyperbolic
conservation laws, \clawpack/\manyclaw,
that implements threaded parallelism for a single compute patch
on the other.  All topological operations on the mesh are handled by \pforest,
while all numerical operations are handled by the \clawpack implementation.
This combination naturally enables a hybrid MPI/threaded parallelization.
We demonstrate scalability of this approach in various ways, where the MPI-only variant
scales up to 16,384 codes

>>>>

We have presented the integration of an MPI-based forest-of-octrees adaptive
meshing code, \pforest, with a numerical solver for hyperbolic
conservation laws, \clawpack/\manyclaw,
that implements threaded parallelism for a single compute patch
on the other.  All topological operations on the mesh are handled by \pforest,
while all numerical operations are handled by the \clawpack implementation.
This combination naturally enables a hybrid MPI/threaded parallelization.
We demonstrate scalability of this approach in various ways, where the MPI-only variant
scales weakly up to 16,384 codes, and strong scalability is linear for up to 256 cores
in the MPI-only and 128 in the threaded implementation

### 2013-07-31 docs/parco13/paper.tex
Make figures fit, improvise conclusion

\begin{figure}
\begin{center}
\plotbox{\includegraphics[width=.4\columnwidth]{quadtree}}
\hspace{.05\columnwidth}
\plotbox{\includegraphics[width=.4\columnwidth]{sphere_plot167}}
\end{center}
\caption{Numerical results for solving the advection equation.  Left:  Unit
square (single quadtree) with a twisting velocity field.  Right: Spherical ball
with a rotational velocity field constructed from two mapped quadtrees.  In
both cases the concentration is color-coded with a sharp gradient shown in red.
The adaptive mesh refinement follows the location of the gradient (the patches
are not shown where they become too fine for display).  Here we use $m = 8$
cells per \clawpack patch.}
\label{fig:results2d}
\end{figure}

@@

\begin{figure}
\begin{center}
  \includegraphics[angle=-90,width=.8\columnwidth]{pdf/uni_advance_strong-crop}
  \\[1ex]
  \includegraphics[angle=-90,width=.8\columnwidth]{pdf/uni_exchange_strong-crop}
\end{center}
\caption{Strong scaling of MPI parallelism for advection on a
  uniformly refined mesh.  Top: run time of time integration.  Bottom:
  time spent in local and parallel neighbor exchanges.  The number of
  patches is $2^{2 \ell}$ at level $\ell$.  The flat lines for the
  smaller runs are caused by having fewer patches than MPI processes,
  leaving some of the processors idle.  For each data point we use the
  time required by the slowest processor to verify that the load is
  balanced equally.
%  all processes. \comment{This last
%    sentence is a bit unclear.  What do you mean exactly?}
}
\label{fig:uniscale}
\end{figure}%
We begin our analysis of Example 1 with uniformly-refined mesh experiments run on TACC's
Stampede supercomputer at different levels of resolution.
We use 1 to 16 MPI processes on one 16-core compute node, obtaining a strong
scaling efficiency for the time integration between 97\%--103\% depending on
the resolution (from levels 1 and 8).  Then, we examine processor counts from
16 to 256 using 16 cores per node.  The scaling behavior remains linear for
the time integration itself, and mostly linear for the local and parallel
neighbor exchanges which require about 10\% of the time integration run time;
see \Fig{uniscale}.

\begin{figure}
\begin{center}
  \includegraphics[angle=-90,width=.8\columnwidth]{../figs_misc/omp.eps}
  \\[1ex]
  \includegraphics[angle=-90,width=.8\columnwidth]{../figs)misc/advance_strong_omp.eps}
\end{center}
\caption{Strong scaling of OpenMP parallelism for advection on a
  uniformly refined mesh.  Top: run time of advection Riemann problem kernel in both an optimized inlined interface and a generic function pointer interface via \manyclaw.  Bottom: MPI + OpenMp scaling results of advance kernel as above.
}
\label{fig:ompscale}
\end{figure}%

In \Fig{ompscale}, we scale intranode parallel calculations and in future iterations on accerelrators the \manyclaw set of solvers were compuled with \forestclaw.  \manyclaw provides a generic set of interators for Riemann problems and related algorithms. In this work we show good scaling results across a single node and coupled to the MPI scaling as well.

@@

>>>>

\begin{figure}
\begin{center}
\plotbox{\includegraphics[width=.37\columnwidth]{quadtree}}
\hspace{.05\columnwidth}
\plotbox{\includegraphics[width=.37\columnwidth]{sphere_plot167}}
\end{center}
\caption{Numerical results for solving the advection equation.  Left:  Unit
square (single quadtree) with a twisting velocity field.  Right: Spherical ball
with a rotational velocity field constructed from two mapped quadtrees.  In
both cases the concentration is color-coded with a sharp gradient shown in red.
The adaptive mesh refinement follows the location of the gradient (the patches
are not shown where they become too fine for display).  Here we use $m = 8$
cells per \clawpack patch.}
\label{fig:results2d}
\end{figure}

@@

\begin{figure}
\begin{center}
  \includegraphics[angle=-90,width=.75\columnwidth]{pdf/uni_advance_strong-crop}
  \\[1ex]
  \includegraphics[angle=-90,width=.75\columnwidth]{pdf/uni_exchange_strong-crop}
\end{center}
\caption{Strong scaling of MPI parallelism for advection on a
  uniformly refined mesh.  Top: run time of time integration.  Bottom:
  time spent in local and parallel neighbor exchanges.  The number of
  patches is $2^{2 \ell}$ at level $\ell$.  The flat lines for the
  smaller runs are caused by having fewer patches than MPI processes,
  leaving some of the processors idle.  For each data point we use the
  time required by the slowest processor to verify that the load is
  balanced equally.
%  all processes. \comment{This last
%    sentence is a bit unclear.  What do you mean exactly?}
}
\label{fig:uniscale}
\end{figure}%
We begin our analysis of Example 1 with uniformly-refined mesh experiments run on TACC's
Stampede supercomputer at different levels of resolution.
We use 1 to 16 MPI processes on one 16-core compute node, obtaining a strong
scaling efficiency for the time integration between 97\%--103\% depending on
the resolution (from levels 1 and 8).  Then, we examine processor counts from
16 to 256 using 16 cores per node.  The scaling behavior remains linear for
the time integration itself, and mostly linear for the local and parallel
neighbor exchanges which require about 10\% of the time integration run time;
see \Fig{uniscale}.

\begin{figure}
\begin{center}
%  \includegraphics[angle=-90,width=.8\columnwidth]{../figs_misc/omp}
%  \\[1ex]
  \includegraphics[angle=-90,width=.7\columnwidth]{../figs_misc/advance_strong_omp}
\end{center}
\caption{Strong scaling of OpenMP parallelism for advection on a
  uniformly refined mesh:
  %Top: run time of advection Riemann problem kernel in both an optimized inlined interface and a generic function pointer interface via \manyclaw.  Bottom:
  MPI + OpenMp hybrid parallelization of the time integration.
}
\label{fig:ompscale}
\end{figure}%

In \Fig{ompscale}, we scale multi-node parallel calculations.
%and in future
%iterations on accelerators 
We use the \manyclaw set of patch-based PDE solvers.
%with
%\forestclaw.  
\manyclaw provides a generic set of interators for Riemann
problems and related algorithms.

%$In this work we show good scaling results
%across a single node and coupled to the MPI scaling as well.

@@

\input{conc}
